---
title: "Activity 5 - Mini-competition Explorations (Group-C)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


```{r, message=FALSE, warning=FALSE, include=FALSE}

library(ggplot2)
library(RColorBrewer)
library(rgdal)
library(gridExtra)
library(rgeos)
library(readr)
library(corrplot)
library(xlsx2dfs)
library(lmtest, pos=4)
library(tidyverse)
library(gapminder)
library(gganimate)
library(car)
library(olsrr)
library(GGally)

```

## Reading in the Datafile

```{r, warning=FALSE, message=FALSE}

## Reading in data for producing graphs

# stdnt <- read.csv("./data/allendale-students.csv")
stdnt <- read_csv(here::here("competition-files","data","allendale-students.csv"))

head(stdnt)
dim(stdnt)

stdnt$housing <- as.factor(stdnt$housing)
stdnt$major <- as.factor(stdnt$major)

```

```{r}
skimr::skim(stdnt)
```


```{r}
summary(stdnt)

```

## Identifying the outliers

```{r}
boxplot(stdnt$distance, col = "orange", horizontal = TRUE )
boxplot(stdnt$scholarship, col = "orange", horizontal = TRUE )
boxplot(stdnt$parents, col = "orange", horizontal = TRUE )
boxplot(stdnt$car, col = "orange", horizontal = TRUE )
boxplot(stdnt$debt, col = "orange", horizontal = TRUE )
```
## Normalizing the data

### Dealing with outliers

```{r}
normdata <- stdnt %>% 
  filter(distance < 250) %>% 
  filter(scholarship < 11000) %>% 
  filter(car > 2) %>% 
  filter(debt > 10000)
boxplot(normdata$distance, col = "orange", horizontal = TRUE, main = "Boxplot of distance")
boxplot(normdata$scholarship, col = "orange", horizontal = TRUE, main = "Boxplot of scholarship" )
boxplot(stdnt$parents, col = "orange", horizontal = TRUE, main = "Boxplot of parents")
boxplot(normdata$car, col = "orange", horizontal = TRUE, main = "Boxplot of car" )
boxplot(normdata$debt, col = "orange", horizontal = TRUE, main = "Boxplot of debt")
```


```{r}
hist(normdata$distance)
hist(normdata$scholarship)
hist(normdata$parents)
hist(normdata$car)
hist(normdata$debt)

boxplot(debt~housing, data = normdata, main = "Boxplot of debt by housing type", xlab = "Housing type", ylab = "Debt", col=topo.colors(7))
boxplot(debt~major, data = normdata, main = "Boxplot of debt by major type", xlab = "Major type", ylab = "Debt", col=topo.colors(7))

ggplot(normdata, aes(x = factor(car), y = debt, fill = major, colour = major)) + 
  geom_bar(stat = "identity", position = "dodge") +
    labs(x="Age of car (in years)", y="Student debt (in dollars)") +
      theme_bw() + 
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
          theme(legend.position = c(0.1, 0.85))

ggplot(normdata, aes(x = factor(car), y = scholarship, fill = major, colour = major)) + 
  geom_bar(stat = "identity", position = "dodge") +
    labs(x="Age of car (in years)", y="Scholarship (in dollars)") +
      theme_bw() + 
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
          theme(legend.position = c(0.1, 0.85))
```

```{r}
normdata <- normdata %>% mutate(sqrt_distance = sqrt(distance), sqrt_scholarship = sqrt(scholarship))
```

```{r}
hist(normdata$sqrt_distance)
hist(normdata$sqrt_scholarship)
```


```{r}
normdata <- normdata %>% select(debt,distance, scholarship, parents, car, housing, major, sqrt_distance, sqrt_scholarship)

```


```{r}
normdata %>% select(debt,distance, scholarship, parents, car) %>% ggpairs
normdata %>% select(debt,sqrt_distance, sqrt_scholarship, parents, car) %>% ggpairs
```


```{r}
corr_matrix <- cor(normdata[,c("sqrt_distance", "sqrt_scholarship", "parents", "car", "debt")], use = "everything")
round(corr_matrix, 3)
corrplot.mixed(corr_matrix, lower.col = "black", number.cex = .7, upper = "ellipse")
```

It seems like sqrt_distance has the strongest linear relationship with debt.

```{r}
model.1 <- lm(debt~sqrt_distance, data=normdata)
summary(model.1)
```

Only 14.9% of the variability in debt that is accounted for this model. 

```{r}
model.2 <- lm(debt~sqrt_distance+sqrt_scholarship+parents+car+housing+major, data=normdata)
summary(model.2)
```

In this model, 69.1% of the variability in debt is accounted. and, we noticed that car has the largest p-value and thus least significant. We could remove it from the model and return the analysis. 

```{r}
model.3 <- lm(debt~sqrt_distance+sqrt_scholarship+parents+housing+major, data=normdata)
summary(model.3)
```

Similarly, we could remove housing from the model and return analysis.

```{r}
model.4 <- lm(debt~sqrt_distance+sqrt_scholarship+parents+major, data=normdata)
summary(model.4)
```

Now, let's remove major from the model. 

```{r}
model.5 <- lm(debt~sqrt_distance+sqrt_scholarship+parents, data=normdata)
summary(model.5)
```

Finally, 67.97% of the variability in student debt is accounted for in this multiple linear regression model. 

```{r}
ols_vif_tol(model.5)
```

The VIF score represents that multicollinearity does not exist in this model. Also, These score indicates that the variables are moderately correlated.


## Residual Analysis

```{r}
bptest(debt~sqrt_distance+sqrt_scholarship+parents, varformula = ~ fitted.values(model.5), studentize=TRUE, data=normdata)
plot(model.5)
```

The insignificant p-value indicates that residuals have constant variance.  
